---
title: "RDBMS - Connecting to a MySQL DB"
author: "Niall McGuinness"
date: "December, 2022"
output: 
  html_document:
    code_folding: hide
---

## Clear console on each execute and set chunk default to not show code

```{r setup, include=FALSE}
cat('\014')
#knitr::opts_chunk$set(include = FALSE)
```

## Download and load libraries

```{r include=FALSE}
 if(!require("RMySQL"))
   install.packages("RMySQL")
 if(!require("dplyr"))
   install.packages("dplyr")
library(RMySQL)
library(dplyr)
```

## Establish a connection to the MySQL DB

```{r include=FALSE}
conn = dbConnect(RMySQL::MySQL(),
                            dbname='sql8583447',
                            host='sql8.freesqldatabase.com',
                            port=3306,
                            user='sql8583447',
                            password = rstudioapi::askForPassword("Enter database password"))
```

## Test the remote connection

```{r}
dbListTables(conn)
dbListFields(conn, "Customers")
```

## Use an R chunk to query and store the results

```{r echo=TRUE}
result <-  dbGetQuery(conn, "SELECT * FROM Customers")
#result
```
## Use an SQL chunk to query and store the results

```{sql connection=conn, output.var="customer_data"}
SELECT * FROM Customers
```

## Accessing the results in an R chunk

```{r}
customer_data
```

```{sql connection=conn}
# INSERT INTO Customers VALUES(3, "mary", "rose", "m@dkit.ie", "1980-07-12")
```

```{sql connection=conn, output.var="customer_data"}
SELECT * FROM Customers
```

```{r}
customer_data
```

## Use an SQL chunk to execute a query with no result

```{sql eval=FALSE, connection=conn, include=FALSE}
#DELETE FROM Customers WHERE fname LIKE "%ar%"
```

## Another way to execute a query in an R chunk

```{r}
target_id <- 2
```

```{sql connection=conn, output.var="result_data"}
SELECT * FROM Customers WHERE id = ?target_id 
```

```{r}
result_data
```

## Do NOT forget to disconnect - otherwise we may hit a remote connection limit

```{r}
dbDisconnect(conn)
```



